---
title: "Descriptive Statistics- Tutorial"
author: "Vivek Jason"
format: 
  html:
    theme: simplex
    toc: true
    toc_float: true
editor: visual
---

### COVID-19 Vaccinations and Death in Malaysia

Call the data in and load packages

```{r, include=FALSE}
#load packages
required_packages <- c("tidyverse", "lubridate", "gtsummary", "rstatix", "janitor", "corrr")
not_installed <- required_packages[!(required_packages %in% installed.packages()[ , "Package"])]    
if(length(not_installed)) install.packages(not_installed)                                           
suppressWarnings(lapply(required_packages, require, character.only = TRUE))

#load data
c19_df <- read.csv("https://raw.githubusercontent.com/MoH-Malaysia/covid19-public/main/epidemic/linelist/linelist_deaths.csv")
```

### **Task 1: Descriptive statistics using `tidyverse`**

Question: Compute the summary statistics (count, mean, standard deviation, minimum, and maximum) of age using tidyverse functions.

Steps:

1.  Install and load the tidyverse package.

2.  Filter the dataset to remove missing values in the "age" column (Note there are no missing values in the dataset- the task is simply meant to simulate the code that would be required if there were).

3.  Use the summary functions from `dplyr` to compute the required summary statistics. In this case- count, mean, standard deviation, minimum, and maximum

```{r}
# Step 1
#install.packages("tidyverse")
library(tidyverse)

# Step 2 & 3
summary_age <- c19_df %>% filter(!is.na(age)) %>% 
  summarise(
  count = n(),
  mean = mean(age),
  sd = sd(age),
  min = min(age),
  max = max(age)
)
summary_age
```

### **Task 2: Descriptive statistics using `gtsummary`**

Question: Create a descriptive statistics table for age, male, bid, and malaysian variables using gtsummary.

Steps:

1.  Install and load the `gtsummary` package.

2.  Create a subset of the data with the selected variables (Note: Select any five variables).

3.  Use the `tbl_summary()` function to compute and display the descriptive statistics.

4.  Stratify by any other selected variable.

```{r}
# Step 1
#install.packages("gtsummary")
library(gtsummary)

# Step 2, 3 & 4
df_subset <- c19_df %>% 
  select(age, male, bid, malaysian) %>% 
  tbl_summary(by = malaysian)
```

### 
